---
title: classification
author: Amanda
date: '2019-10-07'
slug: classification
draft: true
categories: []
tags: []
---

```{r echo=FALSE}
library(tidyverse)
library(magrittr)
library(tidytext)
library(rtweet)
library(here)
token <- readRDS(here("twitter_token.rds"))
```

```{r cache=TRUE}
trump <- get_timeline(user = "realDonaldTrump", n=3200, token = token)
biden <- get_timeline(user = "JoeBiden", n=3200, token = token)
warren <- get_timeline(user = "SenWarren", n=3200, token = token)
sanders <- get_timeline(user = "SenSanders", n=3200, token = token) 

data <- bind_rows(trump, biden, warren, sanders) %>% 
  filter(is_retweet == FALSE) %>% 
  select(screen_name, created_at, text, is_quote, favorite_count, hashtags) 

my_stop_words <- c("donald", "trump", "trump's", "joe", "biden", "bidens", "president", "become", "left", "right", "democrats", "republicans", "&amp")
```

EDA 1: Time series of candidates

```{r}
data %<>% mutate(tweet_day = date(created_at)) 

#
cutoff_date <- data %>%
  group_by(screen_name) %>%
  summarise(first = min(tweet_day)) %>% 
  summarise(cutoff = max(first)) 


tweet_cnts <- data %>% 
  filter(tweet_day >= cutoff_date[[1]]) %>%
  group_by(screen_name, tweet_day) %>%
  summarize(cnt = n()) %>% arrange(desc(tweet_day)) 

tweet_cnts %>%
  ggplot(aes(tweet_day, cnt, color = screen_name)) + geom_line()
```
President Trump tweets the most (no surprise), followed by Joe Biden. Biden has the largest variance. Each of his peaks fall on the day after his democratic debtate, summarizing the points that he made in the debate. 



EDA 2: Process the text. Make word comparison cloud

```{r}
library(wordcloud)
data %>% 
  unnest_tokens(token, text) %>%
  select(screen_name, token) %>%
  anti_join(stop_words, by = c("token"="word")) %>%
  group_by(screen_name) %>%
  count(token, sort = TRUE) %>%
  filter(!token %in% c("t.co", "https", "http", "amp")) %>%
  filter(screen_name=="JoeBiden" | screen_name=="realDonaldTrump") %>%
  pivot_wider(
    names_from = screen_name, 
    values_from = n,
    values_fill = list(n = 0)
  ) %>%
  as.data.frame() %>%
  column_to_rownames("token") %>%
  comparison.cloud(title.colors = "black")

```

Do some classification

1. xgboost
2. DL:
  - fully connected NN
  - wide and deep
  - 1d convnet
  - M's published method
  
  

