---
title: 'Introducing the hddtest R package'
author: AmandaRP
date: '2019-05-24'
slug: new-r-package-hddtest
categories: []
tags: []
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
require(magrittr)
require(tidyr)
require(dplyr)
require(stringr)
require(ggplot2)
set.seed(1969)
load(file="~/hddtest/data/twoNewsGroups.rda")
```

I'd like to introduce a package that I wrote recently, `hddtest`, for two sample hypothesis testing of high dimensional discrete data. The package is currently available on [GitHub](https://github.com/AmandaRP/hddtest). It contains methods for both multivariate binary data and multinomial data.  

To illustrate its use, let's look at a dataset included in the package, called `twoNewsGroups`. It contains document term frequency matrices for two of the [20 newgroups](http://qwone.com/~jason/20Newsgroups/). Each matrix has dimension 594 by 16214. The $(i,j)$ entry of each matrix is the count (term frequency) of the $j^{th}$ word in the $i^{th}$ document. The first matrix in the list contains 594 sampled documents from the `rec.sport.baseball` newsgroup. The second contains 594 sampled documents from the `sci.med` newsgroup.

We might like to compare two sets of documents to determine whether or not they come from the same newsgroup. We'll start by sampling two sets of 200 documents from the `sci.med` newsgroup. We'll use this to simulate the null hypothesis being TRUE, that is the two groups come from the same distribution. 

```{r}
library(hddtest)
data(twoNewsGroups)

num_docs <- 200
row_ids <- 1:nrow(twoNewsGroups$sci.med)
group_1 <- sample(row_ids, num_docs)
group_2 <- sample(row_ids[-group_1], num_docs)
```

Next for each of the two groups, sum the 200 term frequency vectors together. They will be the two multinomial vectors that we test. We'll store the result in a list called, `vecs2Test`.

```{r}
vecs2Test <- list(NA, 2)
vecs2Test[[1]] <- twoNewsGroups$sci.med[group_1, ] %>% 
  colSums() %>% 
  matrix(nrow = 1)
vecs2Test[[2]] <- twoNewsGroups$sci.med[group_2, ] %>%
  colSums() %>% 
  matrix(nrow = 1)
```

Now test the null hypothesis that the two vectors come from the same distribution (i.e. the same newsgroup).

```{r}
vecs2Test %>% multinom.test()
```

The above test produced a significant p-value meaning that we would reject the null. However, the difference isn't very interesting. Instead we could have tested that the differences are within some neighborhood, `delta`.

```{r}
vecs2Test %>% multinom.neighborhood.test(delta = 60)
```

How to choose the appropriate `delta`? The answer may come from subject matter expertise about the problem domain. Or you can run a simulation to gain insight. Below we define a simulation function. 

```{r}
simulation <- function(data, null_hyp, delta, reps = 30, num_docs = c(200, 200)){
  
   vecs2Test <- list(matrix(NA, reps, ncol(data[[1]])), matrix(NA, reps, ncol(data[[1]])))
   
   for(i in 1:reps){
     if(null_hyp){
       # Sample two sets of num_docs from the SAME set of documents     
       row_ids <- 1:nrow(data[[2]])
       group_1 <- sample(row_ids, num_docs[1])
       group_2 <- sample(row_ids[-group_1], num_docs[2])
       vecs2Test[[1]][i, ] <- data[[2]][group_1, ] %>% colSums()
       vecs2Test[[2]][i, ] <- data[[2]][group_2, ] %>% colSums()
       
     }else{
       # Sample num_docs from each of the two DIFFERENT sets of documents
       vecs2Test[[1]][i, ] <- data[[1]][sample(1:nrow(data[[1]]), num_docs[1]), ] %>%
                                      colSums()
       vecs2Test[[2]][i,] <- data[[2]][sample(1:nrow(data[[2]]), num_docs[2]), ] %>%
                                      colSums()
     }
   }
   
   #Perform the test:
   result <- vecs2Test %>% multinom.neighborhood.test(delta = delta)
   
} #end simulation function
```

Now run simulation for varying values of `delta` (in the range 1 to 160) testing both the null and alternative hypotheses for 30 replications each. The resulting plot shows once curve for each the 60 simulations (which evaluate the p-value at each value of `delta`.)

```{r}
delta <- 1:160
p.delta.null <- simulation(data = twoNewsGroups, null_hyp = TRUE, delta = delta)$pvalue_delta
p.delta.alt  <- simulation(data = twoNewsGroups, null_hyp = FALSE, delta = delta)$pvalue_delta
```

```{r, fig.width=8, fig.height=3.5, echo = FALSE}
dat <- data.frame(delta=delta, null = t(p.delta.null), alternative = t(p.delta.alt)) %>%
  gather(-delta, key = "replication", value = pvalue) %>% 
  mutate("hypothesis" = str_extract(replication,"\\w{1,}"))

ggplot(dat, aes(x = delta, y = pvalue, group = replication, colour = hypothesis)) + 
  geom_line() +
  geom_vline(xintercept = 60, linetype = "dotted") +
  annotate("text", x = 80, y = .75, label = "delta = 60") +
  labs(y = "p-value", title = "P-Value Curves for Simulation", color = "Hypothesis") + 
  geom_curve(aes(x = 75, y = .71, xend = 61, yend = .6), curvature = -0.3, arrow = arrow(length=unit(2,"mm")), color = "darkgrey") +
  theme_classic() +
  theme(title = element_text(face = "bold", size = 14),
        panel.grid = element_blank())

```


Notice in the plot above that for `delta` in the range of 50 to 100, the p-value is large when the null hypothesis is true (blue curves) and small when the null hypothesis is false (red curves). So, our previous choice of `delta`=60 seems reasonable.

The methods for this package were developed in collaboration with my UMBC PhD thesis advisor, [Dr Junyong Park](http://www.math.umbc.edu/~junpark/), and documented in [1](https://link.springer.com/article/10.1007%2Fs11749-018-0600-8) and [2](https://www.tandfonline.com/doi/abs/10.1080/03610926.2016.1260743). To see the full list of functions available in `hddtest`, see the [GitHub README](https://github.com/AmandaRP/hddtest). Thanks for reading! 


