---
title: 'Shiny App: Twitter Likes'
author: Amanda Peterson-Plunkett
date: '2019-08-02'
slug: rstats-love
categories:
  - R
tags: []
---




I love that Twitter recently came out with bookmarks! My previous MO on Twitter was to `r emo::ji("heart")` posts that contain information that I want to keep for later, most of which are #rstats tweets. I now have 700+ likes, which are time consuming to browse through when I need to find something. I didn't find an easy way to search likes within Twitter itself, so I built a shiny app to do that using the [`rtweet`](https://rtweet.info/index.html) and [`shiny`](https://shiny.rstudio.com/) R packages. 


# `rtweet` Package

<center>
```{r echo=FALSE}
blogdown::shortcode('tweet', '1109764129005420544')
```
</center>


To grab the tweets, there is a little setup with Twitter that is necessary. The setup is described on the [`rtweet` website](https://rtweet.info/index.html).

```{r echo=FALSE, message=FALSE}
## load rtweet package
library(rtweet)
library(here)
library(tidyverse)
token <- readRDS(str_c(here(),"/twitter_token.rds"))
```

During setup, I ran into a bit of an issue authenticating my Twitter app. I'm working in RStudio on an AWS instance (via a Chromebook), which has a known issue. The issue and its work-around are described by the package author [here](https://github.com/mkearney/rtweet/issues/156). After that speed bump was crossed, the `rtweet` package was extremely easy to use!

# My Twitter Likes

To get a Twitter user's favorites (aka likes), use the `get_favorites` function. Below I'm going to grab  my most recent five favorites:

```{r}
favorites <- get_favorites("DrAmandaRP", n = 5, token = token)
```

The resulting [tibble](https://tibble.tidyverse.org/) is has 91 columns! 


For fun, here is a wordcloud composed of all of my Twitter likes (created using the [`tidytext`](https://cran.r-project.org/web/packages/tidytext/index.html) and [`wordcloud2`](https://cran.r-project.org/web/packages/wordcloud2/index.html) packages):

```{r message=FALSE}
library(wordcloud2)
library(tidytext)
favorites <- get_favorites("DrAmandaRP", n = 1000, token = token)
favorites %>% 
  select(text) %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words) %>%
  filter(!word %in% c("t.co", "https", "http")) %>%
  wordcloud2()

```


# shiny app for searching 

I zeroed in on the following fields: `status_id`, `created_at`, `text`, `hashtags`, `name`, `screen_name`. I put them in a table in a shiny app for easy browsing. Check it out [here](https://amandarp.shinyapps.io/TwitterLikes/)! You can replace my twitter handle with yours (or any Twitter user). It's nothing fancy, but I think it's going to come in handy. You can find my code for the shiny app [on GitHub](https://github.com/AmandaRP/TwitterBrowser/blob/master/TwitterBrowser/app.R).

It would be nice to add functionality to search Twitter bookmarks, but apparently at the time of this post, reading bookmarks isn't yet available in the Twitter API. I'm keeping an eye on [this `rtweet` issue](https://github.com/mkearney/rtweet/issues/344) for updates. 

Have comments? Message me on Twitter: @DrAmandaRP
